Data preparation for estimation of storage models
========================================================

Pfaffenzeller data
------------------

Commodity price data are Grilly and Yang (1988) prices updated by Pfaffenzeller Newbold, and Rayner (2007) and available on [Stephan Pfaffenzeller's personal website](http://www.stephan-pfaffenzeller.com/cpi.html).

US CPI data comes from [Handbook of Labor Statistics, U.S. Department of Labor, Bureau of Labor Statistics](http://www.minneapolisfed.org/community_education/teacher/calc/hist1800.cfm)

```{r}
gycpi <- read.csv('gycpi-2011-01.csv',
                  nrows = 112,
                  row.names = "Year")

uscpi <- read.csv('uscpi-1800-2012.csv',row.names = "Year")

ComPrices <- merge(gycpi,uscpi,by = "row.names")
rownames(ComPrices) <- ComPrices[,1]
ComPrices <- ComPrices[,-1]

ComPrices <- ComPrices[,1:24]/ComPrices[,"USCPI"]

# Commodity in alphabetic order
ComPrices <- ComPrices[,order(colnames(ComPrices))]
```

Define function that calculate descriptive statistics

```{r tidy=FALSE}
require(moments)
StatDes <- function(data) {
  X <- rbind(colMeans(data),
             sapply(data,sd)/colMeans(data),
             skewness(data),
             kurtosis(data)-3,
             sapply(data,function(x) acf(x,plot=FALSE)[[1]][2:3]),
             sapply(data,min),
             sapply(data,max))
  rownames(X) <- c("Mean","CV","Skewness","ExcessKurtosis","AC1","AC2","Min","Max")
  return(t(X))
}
```

### Descriptive statistics for commodity price index 1900-2011

```{r results='asis',echo=FALSE}
library(xtable)
print.xtable(xtable(StatDes(ComPrices)),type='html')
```

### Descriptive statistics for commodity price index 1900-1987
```{r results='asis',echo=FALSE}
print.xtable(xtable(StatDes(ComPrices[as.character(1900:1987),])),type='html')
```


Output commodity price index in a csv file for MATLAB

```{r}
write.csv(ComPrices,file = "ComPrices-1900-2011.csv",quote=FALSE)
```

Deaton and Laroque data
-----------------------

Data provided in appendix of Deaton and Laroque (1995):

```{r tidy=FALSE}
Banana <- c(t(read.table("DL1995/BANA1900.DTA")))
Cocoa <- c(t(read.table("DL1995/COCO1900.DTA")))
Coffee <- c(t(read.table("DL1995/COFF1900.DTA")))
Copper <- c(t(read.table("DL1995/COPP1900.DTA")))
Cotton <- c(t(read.table("DL1995/COTT1900.DTA")))
Jute <- c(t(read.table("DL1995/JUTE1900.DTA")))
Maize <- c(t(read.table("DL1995/MAIZ1900.DTA")))
Palmoil <- c(t(read.table("DL1995/PALM1900.DTA")))
Rice <- c(t(read.table("DL1995/RICE1900.DTA")))
Sugar <- c(t(read.table("DL1995/SUGA1900.DTA")))
Tea <- c(t(read.table("DL1995/TEA1900.DTA")))
Tin <- c(t(read.table("DL1995/TIN1900.DTA")))
Wheat <- c(t(read.table("DL1995/WHEA1900.DTA")))
uscpiDL <- c(t(read.table("DL1995/CPI1900.DTA")))
DL1995Prices <- data.frame(Banana,Cocoa,Coffee,Copper,Cotton,Jute,Maize,
                           Palmoil,Rice,Sugar,Tea,Tin,Wheat,uscpiDL,
                           row.names = 1900:1987)
DL1995RealPrices <- DL1995Prices[,-14]/DL1995Prices[,"uscpiDL"]
```

### Descriptive statistics for Deaton and Laroque data

```{r results='asis',echo=FALSE}
print.xtable(xtable(StatDes(DL1995RealPrices)),type='html')
```

Output commodity price index in a csv file for MATLAB

```{r}
write.csv(DL1995RealPrices,file = "ComPrices-DL1995.csv",
          quote=FALSE)
```

Differences between the two datasets
------------------------------------

It seems that there are discrepancies between the Deaton and Laroque datasets and the Pfaffenzeller ones:

```{r}
gyDL1995 <- gycpi[as.character(1900:1987),colnames(DL1995Prices[,-14])]
diffGY <- DL1995Prices[,-14]-gyDL1995
```


```{r results='asis',echo=FALSE}
print.xtable(xtable(diffGY[rowSums(abs(diffGY)>1e-10)>0,],
                    digits=4,
                    caption="Differences between the two datasets"),
             type='html',caption.placement="top")
```

There are minor differences before 1987. They seem to come mostly from the fact that in Pfaffenzeller's dataset there are less numbers after the decimal points.

1986 is the last year of Grilli and Yang prices. For 1987, price indices may have been constructed differently in the two datasets.

Figures
-------

```{r tidy=FALSE,fig.path='../Figures/RealPrices'}
for(com in colnames(DL1995RealPrices)) {
  plot(1900:1987,
       DL1995RealPrices[,com],
       type="l",
       xlab="Year",
       ylab="Price index",
       main=com,
       las = 1,
       col = "blue")
}
```


References
----------

Deaton and Laroque (1995), Estimating a Nonlinear Rational Expectations Commodity Price Model with Unobservable State Variables, _Journal of Applied Econometrics_ 10, S9--S40.

Grilli and Yang (1988), Primary Commodity Prices, Manufactured Goods Prices, and the Terms of Trade of Developing Countries: What the Long Run Shows, _World Bank Economic Review_ 2(1), 1--47.

Pfaffenzeller, Newbold and Rayner (2007), A Short Note on Updating the Grilli and Yang Commodity Price Index, _World Bank Economic Review_ 21(1), 151--163.